---
title: "WhitePaper"
output: github_document
---
# Abstract


Portland, as countless articles will tell you, has been in the past decade experiencing large amounts of population growth. Migration to and from portland is bound to have economic and cultural effects on the city and its residents. Aditionally, it is worth wondering where people are migrating from and what their characteristics are. Our paper is essentially broken down into three components that relate to residential change in Portland. First, we describe migration, with a focus on migration from the Bay Area and Puget Sound/Seattle Area. Second we dicuss economic disparity in the Portland area, and what factors may play a role in that disparity. Third, we dicuss housing prices and changes in housing prices over the last decade in Portland. It seems that there is a high likelihood that these three components are related. We will attempt to show how they may be related. 


# Migration Over Time
This project has focused primarily on the storyline of the migration of people from the Bay Area and Seattle to Portland. We have found, using IRS data, that the largest sources of migration across time to Portland com from these areas as well as the states of California and Washington. Our data tells us about county to county migration. We can observe the largest contributors over time by viewing the Sankey plot shown below.
```{r 1.1,echo = FALSE, warning = FALSE, error=FALSE}
setwd("~/sd-pdx-sea/Sankey_Run")
source("Refined_Sankey.R")
```
```{r 1.2, fig.wid = 20, fig.height=10}
riverplot(migriver, add_mid_points = FALSE
        ,default_style = styles0, srt = 0, xscale = 1, nodewidth = 1.2)
```

# Economic Disparity


# Housing Prices


#Modelling the relationships